On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-01-08 manuscript.Rmd
analysis of gmin vs hydraulic traits
co-author comments and edits incorporated; main script added; various formatting and other changes

\
Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Nicolas Martin-StPaul^2^, Belinda Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
^2^ INRA

@@

```{r include=FALSE}
includefigs <- T
library(png)
library(grid)
library(knitr)
library(pander)

@@

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
(*expand*) We thus have several different definitions of what constitutes a 'minimum' conductance of leaves. In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance. We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different.

\
(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models. We aim to improve the connection somewhat, and point at obvious gaps in understanding.

@@

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle.         |
| g~min~     | Minimum conductance measured from weight loss of detached leaves. |
| g~dark~    | Night-time conductance, or conductance after significant dark adaptation.              |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~ | Cuvette measurements of g~s~ when photosynthesis rates are very low. |

\
Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001] (g~cut,isol~), or via gas exchange techniques, or mass loss of detached leaves vby sealing the side with stomata (g~cut,seal~). The method we here refer to as minimum conductance is that of mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. The g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

@@

\
(*expand*) The results are presented in Fig. 1. Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton. However, @cavender-bares2007 gdark similar to gmin in oaks. Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their  Fig. A2).

@@

**Figure 1**. Comparison of various estimates of the minimum conductance. Error bars are 95/% confidence intervals. See Table 1 for the various definitions. Different letters denote significant differences (alpha = 0.05). Note the logarithmic scale.

**In case you are interested, this is what the same graph looks like with no log scale**

```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure1_unlog.png")
```

**Figure 1 Version 2**. As Figure 1, but natural scale.

@@

@@

\
It is also telling that there are few reports on the plasticity (intra- or interspecific) of g~0~, perhaps because it is so poorly estimated. Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves. Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the slope was unaffected).

@@

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of Martin-StPaul (cite). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2017. In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because environmental conditions varied little. We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation the ball-berry coefficient (i.e. right-hand side of Eq. 1) is lower (Fig. 3c).

@@

**Figure 3**. (*Might go in SuppInfo?*) Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model. In panel (c) coefficient of variation and R2.

@@

and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1. (*expand*, cite some papers showing these are the responses we expect).

@@

\
Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. Various new models that aim to simulate desiccation have a g~min~ term in it (even when not explicitly defined in the main equations) [@xu2016] (sperry, others), but it is not yet well constrained by data.

@@

Refined by @blackman2016, incorporating more well-known hydraulics traits into the framework.

Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics.

**Could we have a figure showing time to mortality as a function of gmin? I know Nico wants to use this in a separate paper, but it might draw some attention here.**

@@

\
Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance (Fig. 1), in line with previous comparisons [@kerstiens1996]. The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster). A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.

@@

\
It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].

@@

\
(*expand*) Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata (see Section XXX). Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.

@@

**Figure 5**. Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale). Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95/% confidence intervals. Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species. Right panel, data for crops only where each point is a genotype. Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species. References for panels A and B can be found in the SuppInfo 1, references for panel C are XXX. (*have to add some labels*).

@@

(*expand*) From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a decrease in g~min~ (see Table 2 for a summary of the literature); the only exception was @vilagrosa2003 who found no response. A number of studies have attempted to attribute this decrease in g~min~ to a change in cuticular composition. @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes. @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts). @bi2017 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way. @kottapalli2009 higher epicuticular wax load in a drought-tolerant vs. susceptible peanut genotype.

@@

To further study the acclimatory and instantaneous response to temperature, we carried out a case study on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Our results showed a large acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber) (Fig. X). (*expand stats and description of results*).

@@

# Appendix or SuppInfo or Delete

\
To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq. 1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,

\
(2)
$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$

\
where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,

\
(3)
$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$

\
where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies. The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002]. For all simulations with this model, we used the implementation in the *plantecophys* R package [@duursma2015].

>>>>

\
Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
^2^ URFM, INRA, 84000, Avignon, France
^3^ Université Clermont-Auvergne, INRA, PIAF, 63000, Clermont-Ferrand, France

@@

```{r include=FALSE}
includefigs <- FALSE
library(png)
library(grid)
library(knitr)
library(pander)

@@

# Notes

@fernandez2017 : more discussion of cuticular composition, increase of gmin with altitude ('unripe' cuticles due to short growing season), foliar uptake of water (and adaptation of the cuticle to facilitate that)

Relationship between gmin and yield (to mention in the text???):
Clarke et al. 1989 . Genotypes with low gmin yielded 40-150% more in 4 out of 5 driest enviroments in 9 years of study.
Clarke and McCaig (1982 ) low order but significant positive correlation between yield and water retention (low gmin) in genotypes of Triticum

Perhaps include here changes with ontogeny is species with different juvenile and adult leaves?
Climent et al. 2006  and Pardos et al. 2009 - juvenile needles in some pines have higher rate of water loss than adult needles of the same individuals. Darrow et al. 2002  found also that in some New Zealand heteroblastic sp juveniles had a higher gmin than adults but other sp no significant differences.

@hanson2016 large effect of small gmin on estimates of mesophyll conductance, Ci. This is important but not the focus of this review.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
(*expand*) We thus have several different definitions of what constitutes a 'minimum' conductance of leaves. In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance. We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].

\
(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models We aim to improve the connection somewhat, and point at obvious gaps in understanding.

@@

| Variable      |  Definition                                                                                       |
| ------------- | ------------------------------------------------------------------------------------------------- |
| g~cut~        | Conductance of an isolated non-stomatal cuticle.                                                  |
| g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
| g~dark~       | Night-time conductance, or conductance after significant dark adaptation.                         |
| g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~    | Cuvette measurements of g~s~ when photosynthesis rates are very low.                              |

\
Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata (g~cut,seal~) [@kerstiens1996]. The method we here refer to as minimum conductance is that of mass loss of intact detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. Finally, the g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

@@

\
(*expand*) The results are presented in Fig. 1. Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in cotton.

However, @cavender-bares2007 gdark similar to gmin in oaks.

Actual measured stomatal conductance during low photosynthesis (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their Fig. A2).

@@

**Figure 1**. Comparison of various estimates of the minimum conductance. Error bars are 95\% confidence intervals. Different letters denote significant differences (at $\alpha$ = 0.05). See Table 1 for definitions of minimum conductance, and text for the data sources.

@@

- Measurement error, small fluxes

@@

\
It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated. Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves. Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the slope was unaffected).

@@

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry). In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because environmental conditions varied little. We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq. 1 is lower (Fig. 3c).

@@

**Figure 3**. Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model. In panel (c) coefficient of variation and R^2^.

@@

and using the diffusion constraint $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

@@

\
Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. This is particularly important given the limited variation in stomatal response to drought among plants @martinstpaul2017.

@@

More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation.

@burghardt2006 reported on a correlation between g~min~ and the survival time of leaves.

Finally, Martin-StPaul et al 2017 demonstrated in a whole plant model of hydraulic failure that gmin was probably one of the key parameters to explain water potential drops below cavitation thresholds during drought given the relatively conserved range of stomatal closure to drought.  Relying on the Sureau model presented in @martinstpaul2017 we illustrate the critical role of gmin in defining desiccation time.

@@

\
Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance (Fig. 1), in line with previous comparisons [@kerstiens1996]. The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. Boyer et al. 1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases (Boyer et al., 1997)

\
Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster). A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.

@@

\
It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].

@@

\
(*expand*) Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata (see Section XXX). Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions. Also @hanson2016 showing the effect of this on calculation of Ci, and calculation of gmeso.

@@

**Figure 5**. Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale). Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95/% confidence intervals. Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species. Right panel, data for crops only where each point is a genotype. Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species.

@@

(*expand*) From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a decrease in g~min~ (see Table 2 for a summary of the literature); the only exception was @vilagrosa2003 who found no response. A number of studies have attempted to attribute this decrease in g~min~ to a change in cuticular composition. @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes. However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability. @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts). @bi2017 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way. @kottapalli2009 higher epicuticular wax load in a drought-tolerant vs. susceptible peanut genotype.

@@

@@

- Also: there is little information on water loss rate of non-leaf surfaces [@oren2001], should be focus of more research

# SuppInfo

E. parramattensis:
We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively. Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%. Inbuilt fans in each chamber blew air gently across the leaf surfaces. Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour. For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~ following @sinclair1986.

Hakeas:
After 8 months of water stress treatment, the epidermal conductance of detached leaves of three plants per species and treatment was estimated by using the rate of water loss. Leaves were collected in the early morning and subsequently recut under distilled water, rehydrated in darkness for 6 h at 4 °C and scanned to determine leaf area. Leaves were then moved to a controlled environment chamber set at 25 °C, 50% relative humidity and low light (~200 umol) and weighed using a precision balance. They were attached to a ‘clothesline’ and allowed to dehydrated slowly. Leaves were weighted every 15 min for 3 h and then every hour for 3h more. For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and gmin from the slope of the linear part of the curve after stomatal closure, normalized by 2-sided leaf area and adjusted for temperature and humidity within the chamber.
